### 2014-12-12 montana/readme.Rmd
new readme for subdirectories

---
title: "READ ME"
author: "Andrew Tredennick"
date: "December 11, 2014"
output: html_document
---

These are the base folders for the Montana dataset analysis.

### Folder descriptions
* `ipm`: contains subdirectories for vital rate statistical models based on individual level demographic data and code to run single-species integral projection models.

* `quadBM`: contains subdirectories for vital rate statistical models based on plot level data and code to run the quadrat-based population model

* `speciesData`: contains subdirectories for each analysed species (the dominant species) and the average quadrat cover for each species. Species directories hold raw data on individual cover through time in each quadrat

* `weather`: contains the climate file with pre-aggregated variables