### 2015-07-07 manuscript/appA.Rmd
adds appendix A draft

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---

Appendix A: Detailed model descriptions
=======================================
### Tredennick and Adler

Here we describe the statistical models in complete detail, including full description of the Bayesian models and Stan code for each model. We write each model as three parts: (1) a data model (likelihood), (2) a process model, and (3) a parameter model. We then combine these three models to write the posterior and joint distributions as a single model statement. Following the full model expression we provide the Stan code that directly corresponds the statement of posterior and joint distributions. In our expressions of the posterior and joint distributions we use general probability notation where [.] represents some unspecified distribution. We do this for clarity of model presentation. For each model we include a table of prior distributions for necessary parameters. Also for clarity, we avoid subscripting species (the *j* subscripts in the main text).

\newpage{}

Individual level models
-----------------------
### Survival
Data, process, and parameter models.

\begin{align}
&\mathtt{data} \qquad &\textbf{y} \sim \text{Bernoulli}(\boldsymbol{\mu}) \\
&\mathtt{process} \qquad &\text{logit}(\mu_{iQ,t}) = \gamma_{t} + \phi_{Q} + \beta_{t}x_{ij,t} + \omega w_{i,t} + \theta_{k}C_{k,t} \\
&\mathtt{parameters} \qquad &\gamma_{t} \sim \text{Normal}(\tilde{\gamma}, \sigma_\gamma^2)\\
& \qquad & \phi_{Q} \sim \text{Normal}(0,\sigma_Q^2) \\
& \qquad & \beta_{t} \sim \text{Normal}(\tilde{\beta}, \sigma_\beta^2) \\
& \qquad & \omega \sim \text{Normal}(0, 0.0001) \\
& \qquad & \boldsymbol{\theta} \sim \text{Normal}(0, 0.0001)
\end{align}

Full expression of posterior and joint distributions.
\begin{align}
[\boldsymbol{\gamma}, \tilde{\gamma}, \boldsymbol{\phi}, \boldsymbol{\beta}, \tilde{\beta}, \omega, \boldsymbol{\theta}, \sigma_Q^2, \sigma_\beta^2] &\propto \\
&\prod_{t=1}^T \prod_{i=1}^n [y_{iQ,t} | \gamma_{t}, \phi_{Q}, \beta_{t}, \omega, \boldsymbol{\theta}] [\gamma_{t} | \tilde{\gamma}] [\beta_{t} | \tilde{\beta}, \sigma_\beta^2] \times \\
&\prod_{Q=1}^{Q_{tot}} [\phi_{Q} | \sigma_Q^2] \times \\
&[\boldsymbol{\theta}] [\omega] [\tilde{\gamma}] [\tilde{\beta}] [\sigma_Q^2] [\sigma_\beta^2]
\end{align}

Stan code for model.
```{r survival_stan, echo=TRUE, eval=FALSE}
data{
  int<lower=0> N; // observations
  int<lower=0> Yrs; // years
  int<lower=0> yid[N]; // year id
  int<lower=0> Covs; // climate covariates
  int<lower=0> G; // groups
  int<lower=0> gid[N]; // group id
  int<lower=0,upper=1> Y[N]; // observation vector
  matrix[N,Covs] C; // climate matrix
  vector[N] X; // size vector
  matrix[N,2] W; // crowding matrix
}
parameters{
  real a_mu;
  vector[Yrs] a;
  real b1_mu;
  vector[Yrs] b1;
  vector[Covs] b2;
  vector[2] w;
  vector[G] gint;
  real<lower=0> sig_a;
  real<lower=0> sig_b1;
  real<lower=0> sig_G;
}
transformed parameters{
  real mu[N];
  vector[N] climEff;
  vector[N] crowdEff;
  climEff <- C*b2;
  crowdEff <- W*w;
  for(n in 1:N){
    mu[n] <- inv_logit(a[yid[n]] + gint[gid[n]] + b1[yid[n]]*X[n] + 
                         crowdEff[n] + climEff[n]);
  }
}
model{
  // Priors
  a_mu ~ uniform(-300,300);
  w ~ uniform(-100,100);
  b1_mu ~ uniform(-100,100);
  sig_a ~ cauchy(0,5);
  sig_b1 ~ cauchy(0,5);
  sig_G ~ cauchy(0,5);
  gint ~ normal(0, sig_G);
  b2 ~ uniform(-10,10);
  a ~ normal(a_mu, sig_a);
  b1 ~ normal(b1_mu, sig_b1);

  // Likelihood
  Y ~ binomial(1,mu);
}
```