### 2015-07-15 manuscript/appC.Rmd
updates colors in appC figure

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{C\arabic{figure}}

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "appC"
opts_chunk$set(fig.path = paste("appC/figure/", basename, "-", sep=""),
               cache.path = paste("appC/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

Appendix C: Effect of recruitment on IPM uncertainty
=====================================================
In Fig. 4 of the main text where we show the uncertainty around our models' forecasts to a 1% change in climate covariates, the obvious outlier is *Pascopyrum smithii* (PASM). The uncertainty around PASM forecasts from the integral projection model is orders of magnitude larger than for any other species, and is even larger than the uncertainty around the quad-based model forecasts. It turns out that the recruitment parameters we estimate lead to overcompensating density-dependence that is amplified when those parameters are allowed to vary. When we allow all parameters to vary except those associated with recruitment, uncertainty around PASM decreases to levels similar to those associated with other species (Fig. C1). So detailed demographic data is not a cure-all, and can sometimes lead to interesting dynamics when parameter uncertainty is considered.

```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))

```

```{r plot_it, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE, fig.width=5, fig.cap="Equilibrium cover and 90% quantiles around the mean prediction from the IPM when model error and parameter uncertainty are propogated through the simulation phase for all vital rates except recruitment.", dependson="plot-options"}
library(plyr)
library(ggplot2)
setwd("~/Repos/MicroMesoForecast/analysis/")
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, model=NA)
##  IPM 
path <- "./ipm/simulations/results/cc_rectest/"
ipmfiles_mean <- list.files(path)
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
for(file in ipmfiles_mean){
  tmp <- read.csv(paste(path,file,sep=""))
  tmp$model <- "IPM"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
}
meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]

newd <- data.frame(cover=NA, species=NA, 
                   climsim=NA, model=NA, obscover=NA)
for(dospp in unique(meanparam_ipm$species)){
  tmp <- subset(meanparam_ipm, species==dospp)
  obscover <- subset(tmp, climsim=="obs")["cover"]
  tmp2 <- subset(tmp, climsim!="obs")
  tmp2$obscover <- rep(obscover$cover, times = length(unique(tmp2$climsim)))
  newd <- rbind(newd, tmp2)
}
ipm_meansims <- newd[2:nrow(newd),]
ipm_meansims$propdiff <- with(ipm_meansims, log(cover*100)-log(obscover*100))

meanparam_agg <- ddply(ipm_meansims, .(species, climsim, model), summarise,
                       meandiff = mean(propdiff),
                       hidiff = quantile(propdiff, probs = 0.95),
                       lodiff = quantile(propdiff, probs = 0.05))
dgd = position_dodge(width = 0.6)
ggplot(meanparam_agg, aes(color=species))+
  geom_errorbar(aes(x=climsim, ymax=hidiff, ymin=lodiff),
                position=dgd, width=0.25)+
  geom_point(aes(x=climsim, y=meandiff),
                position=dgd, width=0.25)+
  theme_bw()+
#   facet_wrap("species", ncol=1, scales="free")+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Proportional difference")+
  xlab("Climate simulation")+
  scale_color_manual(values=cbPalette[1:4])
```