### 2015-07-31 manuscript/appE.Rmd
adds new results from upping the perturbation percentages for climate

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
bibliography: ~/Dropbox/Bibliography/MicroMeso.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{E\arabic{figure}}

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "appE"
opts_chunk$set(fig.path = paste("appE/figure/", basename, "-", sep=""),
               cache.path = paste("appE/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

Appendix E: Model forecasts under larger climate perturbations
==============================================================
In the main text we show results from model forecasts where we perturbed climate covariates by 1% (Fig. 6). Given the uncertainty in these forecasts, we conducted further simulations where we increased climate covariates by 10%, 20%, and 30% to see how large of a perturbation is necessary to produce forecasts that do not overlap zero. Here we show the results.

```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))

```

```{r figure_e1, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE, fig.width=8.5, fig.height=4, fig.cap="Projected proportional changes in equilibrium cover caused by 10%, 20%, and 30% increases in observed precipitation (+ppt), temperature (+temp), or both (+ppt&temp) as predicted by the individual-level IPM.", dependson="plot-options"}
library(plyr)
library(ggplot2)
library(gridExtra)
####
##  Mean parameter runs
####
setwd("~/Repos/MicroMesoForecast/analysis/ipm/simulations/results/climchange_varyparams/")
meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, pertamount=NA)
##  IPM 
ipmfiles_mean <- list.files()
ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
toget <- grep("10", ipmfiles_mean)
ipmnows <- ipmfiles_mean[toget]
for(file in ipmnows){
  tmp <- read.csv(file)
  tmp$pertamount <- "10%"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "pertamount")])
}
ipm10 <- meanparam_ipm[2:nrow(meanparam_ipm),]

# plot(c(1:2001),subset(ipm20, species=="HECO" & climsim=="temp")["cover"], type="l")

meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, pertamount=NA)
toget <- grep("20", ipmfiles_mean)
ipmnows <- ipmfiles_mean[toget]
for(file in ipmnows){
  tmp <- read.csv(file)
  tmp$pertamount <- "20%"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "pertamount")])
}
ipm20 <- meanparam_ipm[2:nrow(meanparam_ipm),]

meanparam_ipm <- data.frame(cover=NA, species=NA, 
                             climsim=NA, pertamount=NA)
toget <- grep("30", ipmfiles_mean)
ipmnows <- ipmfiles_mean[toget]
for(file in ipmnows){
  tmp <- read.csv(file)
  tmp$pertamount <- "30%"
  meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "pertamount")])
}
ipm30 <- meanparam_ipm[2:nrow(meanparam_ipm),]

pertd <- rbind(ipm10, ipm20, ipm30)
pertd[which(pertd$cover>1),"cover"] <- 1

obsfiles <- ipmfiles_mean[grep("obsClimate",ipmfiles_mean)]
obsd <- data.frame(cover=NA, species=NA, 
                   climsim=NA, pertamount=NA)
for(file in obsfiles){
  tmp <- read.csv(file)
  tmp$pertamount <- "0%"
  obsd <- rbind(obsd, tmp[,c("cover", "species", "climsim", "pertamount")])
}
obsd <- obsd[2:nrow(obsd),]

alld <- rbind(obsd, pertd)

newd <- data.frame(cover=NA, species=NA, 
                   climsim=NA, pertamount=NA, obscover=NA)
for(dospp in unique(alld$species)){
  tmp <- subset(alld, species==dospp)
  obscover <- subset(tmp, climsim=="obs")["cover"]
  tmp2 <- subset(tmp, climsim!="obs")
  tmp2$obscover <- rep(rep(obscover$cover, times = length(unique(tmp2$climsim))), times=3)
  newd <- rbind(newd, tmp2)
}
ipm_meansims <- newd[2:nrow(newd),]
ipm_meansims$propdiff <- with(ipm_meansims, log(cover*100)-log(obscover*100))

meanparam_agg <- ddply(ipm_meansims, .(species, climsim, pertamount), summarise,
                       meandiff = mean(propdiff),
                       hidiff = quantile(propdiff, probs = 0.95),
                       lodiff = quantile(propdiff, probs = 0.05))
dgd = position_dodge(width = 0.9)
meanplot <- ggplot(meanparam_agg)+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_point(aes(x=pertamount, y=meandiff),
             position=dgd, stat="identity")+
  geom_errorbar(aes(x=pertamount, ymax=hidiff, ymin=lodiff),
                position=dgd, width=0.25)+
  theme_bw()+
#   scale_fill_manual(values=c("black", "grey45"), name="PertAmount")+
  facet_wrap(species~climsim, scales="free", nrow=4, ncol=3)+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Mean proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(meanplot)
```