### 2016-04-04 analysis/ipm/vitalRateRegs/regularization/README.Rmd
new README for regularization directory

### README for ridge-regression scripts

*The files in these sub-directories are meant to be run an a High Performance Computing system, NOT on a PC.*

There are directories for each vital rate (growth, survival, and recruitment) that include scripts for running the ridge-regression fits on a High Performance Computing system.

The ``get_plot_lppd.R`` file is script for fetching ridge-regression results (specifically, the log pointwise predictive density from cross-validation model fits) and plotting the lppd versus the prior variance assigned to climate effects.