### 2016-04-04 analysis/ipm/vitalRateRegs/growth/README.Rmd
adds growth regression dir README

### README for growth regression scripts

* The ``*Crowding.csv`` files contain information on the crowding experienced by a focal genet at a specific time.

* There are four ``R`` files:
    + ``crowdingCalculation.R``: calculates the amount crowding experience by a focal genet in a quadrat-year. Results are saved in ``*Crowding.csv`` files to be imported for growth regression fitting script.
    + ``growthAllSpp_STAN.R``: script to read-in and format data, and then fit genet growth regression using Stan (see ``growth.stan`` file for model code).
    + ``import2ipm_means.R``: collates MCMC results and defines function for importing posterior mean model coefficients for IPM simulations.
    + ``import2ipm.R``: collates MCMC results and defines function for importing ranomly selected model coefficients from the full MCMC for IPM simulations with variable parameters.

* The ``growth.stan`` file contains a Stan model for fitting the genet growth regression via ‘No-U-Turn’ Hamiltonian Monte Carlo.

* The ``*.RDS`` files contain the MCMC samples from model fitting for each species.

* The ``cache/`` directory contains old versions of R scripts.