### 2016-04-04 analysis/ipm/vitalRateRegs/recruitment/README.Rmd
adds survival and recreuitment dir READMEs

### README for recruitment regression scripts

* The ``*Crowding.csv`` files contain information on the crowding experienced by a focal genet at a specific time.

* There are four ``R`` files:
    + ``recruitmentAllSpp_STAN.R``: script to read-in and format data, and then fit genet recruitment regression using Stan (see ``recruitment.stan`` file for model code).
    + ``import2ipm_means.R``: collates MCMC results and defines function for importing posterior mean model coefficients for IPM simulations.
    + ``import2ipm.R``: collates MCMC results and defines function for importing ranomly selected model coefficients from the full MCMC for IPM simulations with variable parameters.

* The ``recruitment.stan`` file contains a Stan model for fitting the genet recruitment regression via ‘No-U-Turn’ Hamiltonian Monte Carlo.

* The ``*.RDS`` files contain the MCMC samples from model fitting for each species.