### 2016-04-20 manuscript/components/error_table.tex
updates QBM forecast horizon results and figure; adds error table

\begin{table}[ht]
\centering
\caption{Accuracy (mean absolute error, MAE) and precision (90\% Distance) 
                  of out of sample predictions. Forecasts were made without random 
                  year effects; only climate covariates could explain year-to-year 
                  variation. 90\% Distance refers to the average distance between the 
                  upper and lower 90th percentiles of the 100 predicted values for 
                  each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 4.03 & 24.74 & 9.43 \\ 
  BOGR & QBM & 2.80 & 29.46 & 8.35 \\ 
  HECO & IPM & 0.34 & 1.87 & 1.15 \\ 
  HECO & QBM & 0.29 & 5.53 & 1.18 \\ 
  PASM & IPM & 0.14 & 0.50 & 0.42 \\ 
  PASM & QBM & 0.17 & 1.57 & 0.42 \\ 
  POSE & IPM & 0.45 & 1.63 & 1.25 \\ 
  POSE & QBM & 0.41 & 3.51 & 1.27 \\ 
   \hline
\end{tabular}
\end{table}