### 2016-04-26 manuscript/components/error_table.tex
adds new one step and climate change simulation results

\begin{table}[ht]
\centering
\caption{Accuracy (mean absolute error, MAE) and precision (90\% Distance) 
                  of out of sample predictions. Forecasts were made without random 
                  year effects; only climate covariates could explain year-to-year 
                  variation. 90\% Distance refers to the average distance between the 
                  upper and lower 90th percentiles of the 100 predicted values for 
                  each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 4.61 & 22.41 & 8.26 \\ 
  BOGR & QBM & 4.06 & 29.18 & 8.26 \\ 
  HECO & IPM & 0.59 & 1.93 & 1.22 \\ 
  HECO & QBM & 0.60 & 5.64 & 1.22 \\ 
  PASM & IPM & 0.15 & 0.49 & 0.40 \\ 
  PASM & QBM & 0.20 & 1.59 & 0.40 \\ 
  POSE & IPM & 0.71 & 1.66 & 1.23 \\ 
  POSE & QBM & 0.76 & 3.52 & 1.23 \\ 
   \hline
\end{tabular}
\end{table}