### 2016-10-13 analysis/ipm/simulations/README.Rmd
adds readmes for figshare and versioning

### README for IPM simulation directories

This directory contains R scripts for IPM simulations using the fitted vital rate regressions. There are three simulation scripts:

* ```01a_one-step_ahead_wrapper.R``` -- This script loops through all observed year-to-year transitions,
sets up initial conditions, and then calls the IPM to project the population one time step. The IPM includes climate effects.

* ```01b_one-step_ahead_wrapper_densdep.R``` -- This script loops through all observed year-to-year transitions,
sets up initial conditions, and then calls the IPM to project the population one time step. The IPM does not include climate effects.

* ```02_forecast_horizon_sims.R``` -- Script for IPM forecasts of final observation year from all possible lag years. IPM is initialized with cover at year t-x in quadrat q, and then projected using observed climate to year t. We can then assess the model's forecast horizon and compare it to the IBM.

The three simulation scripts rely on helper functions defined in the remaining .R files. These simulations produce the output summarized in Figures 3 and 4 of the paper.