### 2016-10-13 analysis/quadBM/simulations/README.Rmd

### README for QBM simulation directories

This directory contains R scripts for QBM simulations using the fitted regression. There are three simulation scripts:

* ```01a_loyo_validation_qbm.R``` -- This script loops through all observed year-to-year transitions,
sets up initial conditions, and then calls the QBM to project the population one time step. The QBM includes climate effects.

* ```01b_loyo_validation_qbm_densdep_only.R``` -- This script loops through all observed year-to-year transitions, sets up initial conditions, and then calls the QBM to project the population one time step. The QBM does not include climate effects.

* ```02_forecast_horizon_qbm.R``` -- Script for QBM forecasts of final observation year from all possible lag years. IPM is initialized with cover at year t-x in quadrat q, and then projected using observed climate to year t. We can then assess the model's forecast horizon and compare it to the IBM.

The three simulation scripts rely on helper functions defined in the remaining .R files. These simulations produce the output summarized in Figures 3 and 4 of the paper.