### 2016-10-13 analysis/quadBM/vitalRateRegressions/README.Rmd

### README for vital rate regression Quadrat-Based Model directories

This directory contains sub-directories for fitting the quadrat-based model (``truncNormModel``), a subdirectory containing scripts for running cross-validation and ridge regression for each vital rate (``regularization/``), and a subdirectory containing scripts for leave-one-year-out cross-validation for each vital rate regression (``truncNormModel/validation/``). There is also a ``cache/`` directory containing old versions of scripts.